---
title: "Access to Affordable Finance for Young Agripreneurs"
subtitle: "to Build Youth-Led Agri-Businesses"
author: "Team Data Rangers"
format: 
  html:
    theme: cosmo
    toc: true
    number-sections: true
    toc-title: "Content"
    toc-depth: 2
execute:
  echo: true
  warning: false
  eval: false
---

# Introduction

Access to affordable finance is one of the most pressing barriers limiting youth participation in agriculture in Nigeria. According to the EFInA 2020 survey, only **6%** of youth aged 18–25 can access business loans. With agriculture employing about **35%** of the workforce yet contributing less than **25%** to GDP, it's evident that underinvestment is a challenge.

This project aims to leverage data to **bridge the finance gap** and develop solutions that make agriculture more **bankable and attractive** for youth.

# Objective

The core objective is to develop a **data-driven solution** that enables young agripreneurs to access affordable finance. Our goals include:

- Identifying characteristics that make young farmers creditworthy.
- Predicting repayment likelihood and business success.
- Proposing tools for financiers to evaluate and support youth in agriculture.

# Data Strategy

## Datasets Used

| Dataset Name | Description | Source |
|--------------|-------------|--------|
| [Dataset 1] | [Brief Description] | [Source/Link] |
| [Dataset 2] | [Brief Description] | [Source/Link] |

## Data Collection & Cleaning

- Data imported from [source or format]
- Missing values handled using [method]
- Outliers and anomalies checked using [techniques]
- Feature engineering applied to variables like loan amount, repayment history, crop type, etc.

## Exploratory Data Analysis (EDA)

```{r}
# Example: Visualizing age distribution and access to finance
library(ggplot2)

ggplot(data, aes(x = Age, fill = HasLoanAccess)) +
  geom_histogram(binwidth = 2) +
  labs(title = "Age Distribution and Access to Finance", x = "Age", y = "Count")
```

# Farmer Finance Profile
## Creditworthiness Indicator

```{r}
# Example scoring or correlation heatmap
library(corrplot)
corr_matrix <- cor(data[, c("LoanAmount", "RepaymentRate", "FarmIncome", "TrainingHours")])
corrplot(corr_matrix, method = "circle")

```
Key variables linked to creditworthiness:

Farm size

Previous loan repayment record

Farm income & expenses

Access to extension services or training

## Visualization Dashboard (Optional)
[Embed interactive visualizations using plotly or leaflet]

# Predictive Modeling (Bonus)
## Model Design

```{r}
# Sample logistic regression for predicting loan repayment
model <- glm(RepaymentSuccess ~ Age + FarmSize + Training + Income, data = data, family = "binomial")
summary(model)

```


## Model Performance
Accuracy: XX%

AUC-ROC: XX%

Confusion Matrix:

True Positives: XX

False Positives: XX

etc.

# Solution Proposal
## Concept Overview
We propose a prototype app called "AgriFinanceConnect" which includes:

A loan application portal tailored for young farmers.

A farmer credit scorecard based on our model.

A dashboard for lenders/investors to filter low-risk candidates.

## Prototype Features (Bonus)
Educational Tips: "How to Improve Your Credit Score"

SMS reminders for repayment & training modules

Location-based finance opportunities

## Building Trust
Our solution builds trust by:

Providing transparency through data-backed profiles

Reducing perceived risk for lenders using predictive models

Empowering farmers with knowledge and resources

# Conclusion
By leveraging data analytics, predictive modeling, and user-centered design, this project proposes a scalable, trustworthy pathway to improve financial access for Nigeria’s next generation of agripreneurs.

# References
EFInA. (2020). Access to Financial Services in Nigeria.

World Bank. (2022). Nigeria Employment Statistics.

[Add all other relevant references]
