---
title: "Reducing Post-Harvest Losses "
subtitle: "Empowering Youth-Led Agri-Businesses"
author: "Team Data Rangers"
format: 
  html:
    theme: cosmo
    toc: true
    number-sections: true
    toc-title: "Content"
    toc-depth: 2
execute:
  echo: true
  warning: false
  eval: false
---

# Introduction

Post-harvest losses (PHL) are a critical barrier to food security, economic growth, and youth engagement in Nigeria’s agricultural sector. According to FMARD, **30–50%** of perishable produce is lost annually due to inadequate storage, poor logistics, and weak market systems. This amounts to nearly **\$9 billion in annual losses** (World Bank), equivalent to millions of missed meals and missed business opportunities.

This project explores how **data-driven solutions** can be used to **predict, prevent, and reduce** these losses — turning agriculture into a viable and attractive sector for Nigeria’s youth.

# Objective

To develop an intelligent system that:

-   Identifies **PHL risk factors** (crop type, region, season, method of handling).
-   Predicts **high-loss periods or areas**.
-   Proposes **smart solutions** for storage, distribution, and market access.

# Data Strategy

## Datasets Used

| Dataset Name                  | Description                                        | Source                                 |
|-------------------------------|----------------------------------------------------|----------------------------------------|
| NASS Crop Production Survey   | Production and storage patterns by crop            | National Bureau of Statistics          |
| Climate Data                  | Regional rainfall and temperature patterns         | Nigerian Meteorological Agency (NiMet) |
| FAO Post-Harvest Loss Surveys | Country-specific food loss reports                 | FAO & World Bank                       |
| Custom Survey Data            | Losses by crop, region, storage & logistics method | Team-designed field survey             |

## Data Collection & Cleaning

-   Data imported from \[source or format\]
-   Surveyed farmers in 6 geopolitical zones
-   Integrated geospatial data with market distance, road quality, and temperature
-   Coded crop type, storage duration, and spoilage cause
-   Cleaned nulls, standardized seasonal formats, and geotagged records
-   Missing values handled using \[method\]
-   Outliers and anomalies checked using \[techniques\]
-   Feature engineering applied to variables like loan amount, repayment history, crop type, etc.

## Exploratory Data Analysis (EDA)

```{r}
 # Sample PHL Rate by Crop and Region
library(ggplot2)

ggplot(data, aes(x = Crop, y = LossRate, fill = Region)) +
  geom_col(position = "dodge") +
  labs(title = "Post-Harvest Loss Rate by Crop and Region", x = "Crop", y = "Loss (%)")
```

Key insights:

-Tomatoes and leafy vegetables experience highest losses in southern states.

-Loss rates spike during peak rainy seasons due to poor road access and humidity.

-Regions with fewer storage facilities or long market routes show higher PHL.

# PHL Risk Dashboard

## Interactive Map of High-Risk Zones

\[Use leaflet to display regions with highest recorded PHL by crop and season.\]

```{r}
library(leaflet)

leaflet(data) %>%
  addTiles() %>%
  addCircleMarkers(~longitude, ~latitude,
                   color = ~phl_color, radius = 5,
                   popup = ~paste("Crop:", Crop, "<br>", "Loss Rate:", LossRate, "%"))

```

## Risk Profile Example

+------------------------+------------+------------------------+
| Risk Factor            | Risk Score | Mitigation             |
+========================+============+========================+
| No cold storage nearby | High       | ColdHub Solar Fridges  |
+------------------------+------------+------------------------+
| Peak Rainfall Season   | High       | Mobile Drying Stations |
+------------------------+------------+------------------------+
| > 40Km to Market       | Medium     | Partner with Logistics |
+------------------------+------------+------------------------+

# Predictive Modeling (Bonus)

## Model: Predicting PHL Risk

```{r}
# Random Forest model to predict PHL likelihood
library(randomForest)

phl_model <- randomForest(LossRate ~ Crop + Region + Season + DistanceToMarket + StorageType,
                          data = data, ntree = 100)

print(phl_model)

```

## Performance Summary

-   RMSE: XX%

-   Top Predictors: Crop Type, Season, Storage Access

# Solution Proposal

## Concept: **AgriShield Nigeria**

An integrated digital platform that:

-   **Predicts PHL hotspots** using real-time weather, market, and crop data

-   **Connects farmers to ColdHubs** (solar-powered storage)

-   **Links to logistics via APIs** like **Hello Tractor** or local startups

-   Offers a **mobile dashboard** for young farmers to:

```         
-    Monitor expected spoilage

-    Book cold storage or transport slots

-    View optimal harvest/market times
```

## Real-World Models to Re-Implement

| Real- World Solution    | Function                                          | Re-Implementation                                          |
|-------------------------|---------------------------------------------------|------------------------------------------------------------|
| ColdHubs (Nigeria)      | Solar - powered walk in cold rooms                | Expand via co-op partnerships or youth-owned units         |
| Twiga Foods (Kenya)     | Digital supply chain, logistics & demand matching | Develop a local app connecting smallholders to bulk buyers |
| Hello Tractor (Nigeria) | Tractor sharing platform                          | Extend to cold truck or mobile dryer sharing               |
| Koolboks (Africa)       | Solar freezers with pay-as-you-go                 | Use for high-spoilage zones in fruit & vegetable corridors |
| AgUnity (Global)        | Blockchain recordkeeping for farmers              | Pilot digital ledgers for PHL tracking and transparency    |

## Impact for Youth

-   Unlocks new agri-business models (e.g., storage-as-a-service).

-   Builds digital skills through mobile app management.

-   Increases profitability and encourages retention in the value chain.

## Prototype Features (Bonus)

-   **Smallholder version:** SMS interface to book storage or transport

-   **Investor dashboard:** Highlight PHL investment needs by region

-   **Farmer education popups:** e.g., “Dry maize to 13% moisture before storing”

## Conclusion

By combining predictive analytics, mobile platforms, and logistics innovation, this solution empowers Nigeria’s youth to lead a revolution in agricultural efficiency — reducing waste and turning risk into opportunity.

# References

-   FMARD. (2021). **Post-Harvest Losses in Nigeria**.

<!-- -->

-   World Bank. (2022). **Food Loss and Waste Reports**.

<!-- -->

-   ColdHubs. <https://www.coldhubs.com/>

-   Hello Tractor. <https://www.hellotractor.com/>

-   Twiga Foods. <https://www.twigafoods.com/>

-   FAO. (2020). **The State of Food and Agriculture**.

\[Add all other relevant references\]
