---
title: "Reducing Post-Harvest Losses "
subtitle: "Empowering Youth-Led Agri-Businesses"
author: "Team Data Rangers"
format: 
  html:
    theme: cosmo
    toc: true
    number-sections: true
    toc-title: "Content"
    toc-depth: 2
execute:
  echo: true
  warning: false
  eval: false
---

# Introduction

Post-harvest losses (PHL) are a critical barrier to food security, economic growth, and youth engagement in Nigeria’s agricultural sector. According to FMARD, **30–50%** of perishable produce is lost annually due to inadequate storage, poor logistics, and weak market systems. This amounts to nearly **\$9 billion in annual losses** (World Bank), equivalent to millions of missed meals and missed business opportunities.

This project explores how **data-driven solutions** can be used to **predict, prevent, and reduce** these losses — turning agriculture into a viable and attractive sector for Nigeria’s youth.

# Objective

To develop an intelligent system that:

-   Identifies **PHL risk factors** (crop type, region, season, method of handling).
-   Predicts **high-loss periods or areas**.
-   Proposes **smart solutions** for storage, distribution, and market access.

# Data Strategy

## Datasets Used

This section provides overview of datasets related to post-harvest losses, food prices, logistics, market access, and population dynamics in Nigeria. All datasets are stored in Excel format and compiled from multiple credible sources.

## Dataset Summary Table

+-------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Dataset Name**              | **Description**                                                                                                                              | **Source**                                                                                                                                                |
+===============================+==============================================================================================================================================+===========================================================================================================================================================+
| **Post-Harvest Loss Dataset** | Data on post-harvest loss amounts (in tonnes), percentage of production lost, food price impacts, logistics factors, and production volumes. | FAOSTAT 2017, The Herald, AfricaCheck36, World Bank Open Data, UN Population Prospects, WRI CAIT Climate Data Explorer, FAO et al (2019), WB Calculations |
+-------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Food Price Dataset**        | Data on food price levels, variations by crop, and economic implications of post-harvest losses.                                             | World Bank Open Data, The Herald                                                                                                                          |
+-------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Logistics Dataset**         | Data on storage and transportation capacity (in tonnes) and losses due to poor infrastructure.                                               | AfricaCheck36, FAOSTAT 2017                                                                                                                               |
+-------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Market Dataset**            | Information on market accessibility, regional trade flows, and volumes affected by post-harvest losses.                                      | World Bank Open Data, FAOSTAT                                                                                                                             |
+-------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Population Dataset**        | Data on current population, food insecurity prevalence (44%), and projections (50% increase by 2050).                                        | UN Population Prospects, World Bank Open Data                                                                                                             |
+-------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Other dataset**             | These includes combined metrics based on crop type, population                                                                               | FAO Food Loss and Waste Platform, HarvestChoice SPAM, Harvard Dataverse                                                                                   |
+-------------------------------+----------------------------------------------------------------------------------------------------------------------------------------------+-----------------------------------------------------------------------------------------------------------------------------------------------------------+

## Data Collection & Cleaning

-   Data were downloaded from source and imported
-   Variable names were cleaned and rensme e.g crop type, storage duration, and spoilage cause
-   Cleaned nulls, reshape wide format data with Python and R
-   Missing values were left as missing as they are less than 5% of the data

## Exploratory Data Analysis (EDA)

```{r}
 df_cleaned <- data_list[[2]] %>%
    mutate(
      date = mdy(date), # Convert date to a Date type
      market_id = as.numeric(market_id), # Convert market_id to numeric
      latitude = as.numeric(latitude),   # Convert latitude to numeric
      longitude = as.numeric(longitude),  # Convert longitude to numeric
      `price(NGN)` = as.numeric(`price(NGN)`), # Convert price(NGN) to numeric
      `price(USD)` = as.numeric(`price(USD)`)  # Convert price(USD) to numeric
    )
  
  
  # Major Insight Example for Logistics
  data_list[[2]] %>%
    group_by(State) %>%
    mutate(`price(NGN)` = as.numeric(gsub("[^0-9.]", "", `price(NGN)`))) %>% # Remove non-numeric characters except decimal point
    summarise(Avg_Price = mean(`price(NGN)`, na.rm = TRUE))
```

Key insights:

Analyzing post-harvest losses in Nigeria reveals that states like Kaduna and Katsina face the biggest challenges in preserving their produce. However, it's interesting to see that high losses don't always translate to high market prices. For instance, Benue manages lower losses but isn't a leader in pricing, while Sokoto's unusually low prices and Abia's high ones highlight the complex interplay of factors like market access and demand that truly shape the agricultural value chain beyond just how much is lost after harvest.

# PHL Risk Dashboard

## Interactive Map of High-Risk Zones

\[Use leaflet to display regions with highest recorded PHL by crop and season.\]

```{r}
library(leaflet)

leaflet(data) %>%
  addTiles() %>%
  addCircleMarkers(~longitude, ~latitude,
                   color = ~phl_color, radius = 5,
                   popup = ~paste("Crop:", Crop, "<br>", "Loss Rate:", LossRate, "%"))

```

## Risk Profile Example

+------------------------+----------------+------------------------+
| Risk Factor            | Risk Score     | Mitigation             |
+========================+================+========================+
| No cold storage nearby | High           | ColdHub Solar Fridges  |
+------------------------+----------------+------------------------+
| Peak Rainfall Season   | High           | Mobile Drying Stations |
+------------------------+----------------+------------------------+
| > 40Km to Market       | Medium         | Partner with Logistics |
+------------------------+----------------+------------------------+

# Predictive Modeling (Bonus)

## Model: Predicting PHL Risk

```{r}
# Random Forest model to predict PHL likelihood
library(randomForest)

phl_model <- randomForest(LossRate ~ Crop + Region + Season + DistanceToMarket + StorageType,
                          data = data, ntree = 100)

print(phl_model)

```

## Performance Summary

-   RMSE: XX%

-   Top Predictors: Crop Type, Season, Storage Access

# Solution Proposal

## Concept: **AgriShield Nigeria**

An integrated digital platform that:

-   **Predicts PHL hotspots** using real-time weather, market, and crop data

-   **Connects farmers to ColdHubs** (solar-powered storage)

-   **Links to logistics via APIs** like **Hello Tractor** or local startups

-   Offers a **mobile dashboard** for young farmers to:

```         
-    Monitor expected spoilage

-    Book cold storage or transport slots

-    View optimal harvest/market times
```

## Real-World Models to Re-Implement

+-------------------------+---------------------------------------------------+------------------------------------------------------------+
| Real- World Solution    | Function                                          | Re-Implementation                                          |
+=========================+===================================================+============================================================+
| ColdHubs (Nigeria)      | Solar - powered walk in cold rooms                | Expand via co-op partnerships or youth-owned units         |
+-------------------------+---------------------------------------------------+------------------------------------------------------------+
| Twiga Foods (Kenya)     | Digital supply chain, logistics & demand matching | Develop a local app connecting smallholders to bulk buyers |
+-------------------------+---------------------------------------------------+------------------------------------------------------------+
| Hello Tractor (Nigeria) | Tractor sharing platform                          | Extend to cold truck or mobile dryer sharing               |
+-------------------------+---------------------------------------------------+------------------------------------------------------------+
| Koolboks (Africa)       | Solar freezers with pay-as-you-go                 | Use for high-spoilage zones in fruit & vegetable corridors |
+-------------------------+---------------------------------------------------+------------------------------------------------------------+
| AgUnity (Global)        | Blockchain recordkeeping for farmers              | Pilot digital ledgers for PHL tracking and transparency    |
+-------------------------+---------------------------------------------------+------------------------------------------------------------+

## Impact for Youth

-   Unlocks new agri-business models (e.g., storage-as-a-service).

-   Builds digital skills through mobile app management.

-   Increases profitability and encourages retention in the value chain.

## Prototype Features (Bonus)

-   **Smallholder version:** SMS interface to book storage or transport

-   **Investor dashboard:** Highlight PHL investment needs by region

-   **Farmer education popups:** e.g., “Dry maize to 13% moisture before storing”

## Conclusion

By combining predictive analytics, mobile platforms, and logistics innovation, this solution empowers Nigeria’s youth to lead a revolution in agricultural efficiency — reducing waste and turning risk into opportunity.

# References

-   FMARD. (2021). **Post-Harvest Losses in Nigeria**.

<!-- -->

-   World Bank. (2022). **Food Loss and Waste Reports**.

<!-- -->

-   ColdHubs. <https://www.coldhubs.com/>

-   Hello Tractor. <https://www.hellotractor.com/>

-   Twiga Foods. <https://www.twigafoods.com/>

-   FAO. (2020). **The State of Food and Agriculture**.

\[Add all other relevant references\]
